Biomass and species richness temporal trends drives the assembly and the disassembly of food-web in natural fish assemblages

### 2021-08-25 paper/manuscript.Rmd
Clean appendix

---
title: "Biomass and species richness temporal trends drives the assembly and the disassembly of food-web in natural fish assemblages"
author: Alain Danet, Maud Mouchet, Loubna El Madouri, Elisa Thébault and Colin Fontaine (ordre non fixé)
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

@@

## community structure are dynamics and affect ecosystem functions

Biodiversity affects ecosystem properties such as total biomass, productivity
and stability [@tilman_biodiversity_2014; @loreau_biodiversity_2001]. However,
current biodiversity is at risk, numerous studies reported dramatic declines of
population and community biomass [@ceballos_biological_2017;
@hallmann_more_2017; @albano_native_2021], as well as decrease in species
richness 
[@barnosky_has_2011; @newbold_global_2015]. Consequences of such trends on
various aspects of community structure are not well-known whereas numerous
ecological studies have shown that the different aspects of community structure
are tightly linked [REF].

The temporal trends of species richness and community biomass are indeed
predicted to affect community structure. The positive link between Biodiversity
and ecosystem functioning shown in experimental studies predict that loss and
gain of species richness should respectively result in loss and gain of
community biomass [@tilman_biodiversity_2014; @pennekamp_biodiversity_2018]. But
this relation is not always found in natural settings [@grace_integrative_2016;
@li_relationship_2018; @maureaud_biodiversityecosystem_2019]. Numerous studies
find relation between species richness and evenness, but their disagree on the
sign of the relationships [@soininen_relationship_2012; @ma_species_2005;
@maureaud_biodiversityecosystem_2019], highlighting the diversity of assembly
processes in ecological communities.

The distribution and the number of ecological interactions, i.e. the structure
of ecological networks, also covaries with species richness and community
biomass. In food-webs, connectance, which measures the degree of generalism in a
network, has been theoretically shown to positively scale with
community biomass [@thebault_food-web_2003]. Connectance has been also shown to
be negatively related to species richness in empirical food-webs
[@dunne_network_2006]. Trophic level has also been shown to be positively linked
to total biomass in theoretical models [@wang_biodiversity_2018]. All together,
these studies show that the different dimensions of network structure are 
closely linked.

@@

Using a standardized monitoring of fish freshwater communities on the period
1995-2018, we assessed the relationships among temporal trends of food-web
structure, trends of community biomass and trends in species richness.  
Moreover, we tested if the those relationships differs between communities that
loosed or gained biomass/species richness across time, reflecting potential
difference between disassembly and assembly processes. We finally assessed if
spatial relationships display the same patterns as temporal ones, to assess
the validity of space for time substitution.

@@

@@

@@

Because we had few informations about trophic interactions between resources, we
selected only the fish-fish interactions in subsequent analysis.

@@

$dN = dB + dR$

@@

The significativity of the effects were tested with ANOVA.  In order to be able
to compare the magnitude of the effects of community biomass and species
richness in both temporal and spatial models, we computed standardized slopes
($r_{\delta}$) and their associated standard errors.

@@

tps_bm_sest <- tps_bm$sest
names(tps_bm_sest) <- tps_bm$rep
```
We found a significant positive relationship between the temporal trends of
community biomass from one side and species richness, average trophic level,
connectance from the other side (Fig. \@ref(fig:bm), Table \@ref(tab:aov),
respectively  $r_\delta = `r tps_bm_sest["rich"]`$, $`r tps_bm_sest["tlvl"]`$, 
$`r tps_bm_sest["ct"]`$), meaning that they temporally increased when the community
biomass temporally increased and respectively that they temporally decreased
when the community biomass decreased in time.

We found a significant negative relationship between the temporal trends of
community biomass and that of the pielou of biomass 
( $r_\delta = `r tps_bm_sest["p_bm"]`$), meaning that the eveness of biomass
across species
temporally increased when the community biomass
temporally decreased and respectively that the evenness of biomass temporally
decreased when the community biomass temporally increased (Fig. \@ref(fig:bm),
Table \@ref(tab:aov)). We also found the same pattern between the temporal
trends of community biomass and that of the pielou of individual number
($r_\delta = `r tps_bm_sest["p_n"]`$).

@@

The temporal trends of community biomass and pielou of biomass, had a positive
relationship with that of species richness (Fig. \@ref(fig:rich), Table
\@ref(tab:aov), $r_\delta = `r tps_r_sest["p_bm"]`$) while we did not find
significant relationships between the temporal trends of average trophic level,
connectance, pielou of individual number and the one of species richness (Fig.
\@ref(fig:rich), Table \@ref(tab:aov)).

@@

We found that species richness, connectance, average trophic level were
spatially positively related to community biomass (resp. 
$r_\delta = `r sp_b_sest["rich"]`$, $`r sp_b_sest["ct"]`$, 
$`r sp_b_sest["tlvl"]`$, Fig.
\@ref(fig:bm), Table \@ref(tab:aov-sp)) while the pielou of biomass was
negatively related to community biomass ($r_\delta = `r sp_b_sest["p_bm"]`$).
However, we found no spatial relation between the pielou of individual and
community biomass.

@@

Community biomass, the pielou of biomass and individual were
positively related to species richness (resp. 
$r_\delta = `r sp_r_sest["bm"]`$, $`r sp_r_sest["p_bm"]`$,
$`r sp_r_sest["p_n"]`$) while average
trophic level and connectance were negatively related to
species richness (resp. $r_\delta = `r sp_r_sest["tlvl"]`$, $`r sp_r_sest["ct"]`$,
Fig. \@ref(fig:rich), Table \@ref(tab:aov-sp)).

@@

We found that the temporal increase of the average trophic level was positively
linked to increase of community biomass in line with a previous model suggesting
that the presence of higher predator can lead to higher community biomass
[@wang_biodiversity_2018]. Increase of average trophic level can be associated
to either gain of biomass in the highest trophic level nor losses of biomass in
the lowest trophic. In our case, such gain or losses of average trophic level is
probably associated to the variations of the presence of large fishes. The
finding that the evenness of biomass, measured by the pielou, decreased through
time as community biomass decrease and reversely, meaning that biomass
distribution across species respectively become respectively less and more even
when community biomass decrease and increase. We also found that evenness
decreased temporally of number of individual when community biomass increased
temporally, confirming that the temporal trends of community biomass in our
study is linked to big individuals. It further indicates that temporal loss of
community biomass are linked with the loss of the higher trophic level.

We also found in space that evenness of biomass was positively related to
community while eveness of individual number had no relation. It constrasts with
previous studies which found a negative relationship between biomass and eveness
of individual in marine and plant communities [@drobner_relation_1998;
@maureaud_biodiversityecosystem_2019] or studies that found more even plant
communities in terms of individual number display higher community biomass
[@wilsey_biodiversity_2000].

@@

Surprinsingly, we found that the temporal trends of connectance was not related to
the one of species richness as found spatially in most empirical food-web
[@dunne_network_2006] and also in the present study. This result suggests that
the structure of food-webs might be not very sensitive to variation of species
richness and that species loss and gain concerns species of low trophic level,
which in turn does affect connectance very little. However, the relationship become
slightly negative for species poor communities. An explanation might be that
species-rich communities already contains a significant number of piscivores,
which have the most links whereas species poor communites contains few
piscivores.

@@

This result has profound implications for conservation. As more and more
studies report community biomass and richness trends [@hallmann_more_2017;
@ceballos_biological_2017], it is important to note that they might not be
equivalent in terms of consequences on the structure of the community. Further,
studies should investigate the temporal trends of several dimensions of
community structure simultaneously because it drives many aspects of ecosystem
functioning.

@@

## Conclusion

Our study shows that richness and biomass temporal trends in natural fish
communities relate on different ecological processes, biomass temporal dynamics
being related to variation of highest trophic level while it is not the case for
species richness. Our work also suggests that space for time substitution may be
relevant to predict food-web community structure in case of biomass variation
but not for species richness variation. We look forward to next investigations
to see the generality of assembly rules in food-webs. There are crucial to
better understand the temporal dynamics of food-webs structure and then its
impact on ecosystem functioning.

@@

@@

fig1_2 <- get_plot_rich_bm(
  predict_plot = pred_bm_rich_mono_stable_trends,
  get_list = TRUE,
  rm_legend = FALSE,
  bm_x = "log_bm_std",
  bm_y = c("ct_ff", "w_trph_lvl_avg", "piel_bm"), #, "piel_nind"
  rich_y = c("ct_ff", "w_trph_lvl_avg", "log_bm_std", "piel_bm"), #, "piel_nind"
  rich_x = "log_rich_std"
  )
p_bm <- fig1_2$bm
names(p_bm) <- c("ct_ff", "w_trph_lvl_avg", "piel_bm")
plot_grid(plotlist = p_bm)

@@

```{r}
fig_bm_rich_cap <- paste0(
  "Relationship among the temporal trends of community structure
  and the one of species richness and community biomass. The lines
  surrounding each point represent the standard error associated to the temporal
  trends."
)
```
```{r bmr, fig.height = 12, fig.cap=fig_bm_rich_cap}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
plot_grid(
  plot_grid(plotlist = fig2$bm, nrow = 1),
  plot_grid(plotlist = fig2$rich, nrow = 1),
  nrow = 2
)
```

@@

```{r}
loadd(pred_bm_rich_mono_stable_trends, anova_table_bm_rich_mono_stable_trends)
anova_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
  ) %>%
  kable(.,
    booktabs = TRUE,
    row.names = FALSE,
    caption =
      paste0(
      "ANOVA results for the links between the temporal trends of
      community biomass and the temporal trends of the network structure."
    ),
  label = "aov"
  ) %>%
  kable_styling(latex_options = c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

>>>>

---
title: "Biomass and species richness temporal trends drives the assembly and the disassembly of food-web in natural fish assemblages"
author: Alain Danet, Maud Mouchet, Loubna El Madouri, Elisa Thébault and Colin Fontaine (ordre non fixé)
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

- TODO:
  - [ ] RM Pielou of abundances
  - [ ] Interaction tables 
  - [ ] Interaction plots

@@

## community structure is dynamic and affect ecosystem functions

Biodiversity affects ecosystem properties such as total biomass, productivity
and stability [@tilman_biodiversity_2014; @loreau_biodiversity_2001]. However,
current biodiversity is at risk, numerous studies reported dramatic declines of
population and community biomass [@ceballos_biological_2017;
@hallmann_more_2017; @albano_native_2021], as well as decrease in species
richness [@barnosky_has_2011; @newbold_global_2015]. Whereas numerous
ecological studies have shown that the different aspects of community structure
are tightly linked [@thebault_food-web_2003; @dunne_network_2006;
@soininen_relationship_2012], the consequences of temporal variations of community
biomass and species richness on community structure are not well-known.

## Variations of community biomass and species richness are predicted to affect community structure

Changes of species richness and community biomass are indeed
predicted to affect community structure. The positive link between biodiversity
and ecosystem functioning in experimental studies predict that loss and
gain of species richness should respectively result in loss and gain of
community biomass [@tilman_biodiversity_2014; @pennekamp_biodiversity_2018]. But
this relation is not always found in natural settings [@grace_integrative_2016;
@li_relationship_2018; @maureaud_biodiversityecosystem_2019].

Variations of species richness can also result in variation of evenness, i.e.
equitability of species abundances or biomasses. Several studies shows that
higher evenness of abundances is linked to higher plant biomass in grasslands
and primary productivity in marine ecosystems [@soininen_relationship_2012].
Furthermore, studies reported that higher evenness plays a role of insurance in
face of perturbations [@wittebolle_initial_2009]. Indeed, most studies reported
positive relationship between species richness and evenness, but also neutral
and negative ones [@soininen_relationship_2012; @ma_species_2005;
@maureaud_biodiversityecosystem_2019].

## Variations of community biomass and species richness are predicted to affect food-web structure

The distribution and the number of ecological interactions, i.e. the structure
of ecological networks, also covaries with species richness and community
biomass. In food-webs, connectance, which measures the degree of generalism in a
network, has been theoretically shown to positively scale with community biomass
[@thebault_food-web_2003].  Connectance has been also shown to be negatively
related to species richness in empirical food-webs [@dunne_network_2006].
Trophic level has also been shown to be positively linked to total biomass in
theoretical models [@wang_biodiversity_2018].

The different dimensions of network and community structure are closely linked
but, we have no direct evidence on how they covary according to the
loss or gain of species richness and community biomass across time.

## Predictions come mainly from space-for-time substitution

@@

Using a standardized monitoring of fish freshwater communities on the period
1995-2018, we assessed the relationships among temporal trends of food-web
structure, trends of community biomass and trends in species richness. We
then assessed if spatial relationships display the same patterns as temporal
ones, to assess the validity of space for time substitution.

@@

## Data

@@

## Biomass

We first inferred the body mass of each fish individual thanks to an allometric
law, $B_i = 0.01 \times l_i^3.03$, $B_i$ being biomass in gram and $l_i$ the
body length in millimetre of the individual $i$ [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

@@

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_nodate]: (1) we first computed the metaweb, describing the
potential trophic interactions between all species size classes, (2) then we
extracted each local food-webs by extracting species size classes and their
interactions according their presence in a given sampling event.

We inferred food-web structure according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019; @bonnaffe_comparison_2021;
@danet_species_nodate]. The method is extensively described in
@bonnaffe_comparison_2021 and @danet_species_nodate. The food diet of stream
fishes, as many organisms, change along their life and therefore according to
their body size. Each fish species was divided in nine size classes, named
trophic species, equally distributed in their observed body size range.
A trophic species is a group of individuals of the same species and size class
that are expected to share the same preys and predators. We also added seven
nodes present in the ontogenic food diet database: detritus, biofilm,
phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos. 
The resulting metaweb contained 45 fish species divided into nine size classes
plus 7 resource nodes, summing up to 412 nodes.  Consideration of size classes,
i.e. trophic species instead of species, improves inference of trophic
interactions [@jennings_weak_2001].

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or
three life stages by fish species, each one being delimited by body size
[@bonnaffe_comparison_2021; @danet_species_nodate]. The life stage of a fish
trophic species was determined by the centre of its body size range, hereafter
midpoint. The ontogenic fish diet database was filled according to fishbase
(Froese et al. 2021) and literature [@bonnaffe_comparison_2021;
@danet_species_nodate].

The feeding interactions between fishes were based on the body size ratio
between predators and preys. As this, we defined the predation window of a
piscivorous trophic species as 3% to 45% of its midpoint
[@mittelbach_ontogeny_1998; @claessen_impact_2002; @hart_handbook_2008].  We set
a trohic interaction  A trophic link was defined between the piscivorous trophic
species and all the trophic species whose midpoint in body size range was
included in the predation window.  The fish-resource trophic links were set
between the trophic species and the resources nodes according to the food items
of their size class.  Finally, the feed interactions between resources were
defined according to the literature [@allan_stream_2007; @hart_handbook_2008].

@@

$dN = \alpha + \beta_1dB + \beta_2dR + \beta_3dBdR + \epsilon$

@@

The significativity of the effects were tested with ANOVA. In order to compare
the magnitude of the effects of community biomass and species richness in both
temporal and spatial models, we computed standardized slopes ($r_{\delta}$) and
their associated standard errors.

@@

tps_bm_sest <- tps_bm$sest
names(tps_bm_sest) <- tps_bm$rep
```
We found a significant positive relationship between the temporal trends of
community biomass from one side and species richness, average trophic level,
connectance from the other side (Fig. \@ref(fig:bef) and \@ref(fig:bmr), Table \@ref(tab:aov),
respectively  $r_\delta = `r tps_bm_sest["rich"]`$, $`r tps_bm_sest["tlvl"]`$, 
$`r tps_bm_sest["ct"]`$), meaning that they temporally increased when the community
biomass temporally increased and respectively that they temporally decreased
when the community biomass decreased in time.

We found a significant negative relationship between the temporal trends of
community biomass and that of the pielou of biomass 
( $r_\delta = `r tps_bm_sest["p_bm"]`$), meaning that the eveness of biomass
across species temporally increased when the community biomass temporally
decreased and respectively that the evenness of biomass temporally decreased
when the community biomass temporally increased (Fig. \@ref(fig:bmr),
Table \@ref(tab:aov)).

---
#We also found the same pattern between the temporal
#trends of community biomass and that of the pielou of individual number
#($r_\delta = `r tps_bm_sest["p_n"]`$). 
---

@@

The temporal trends of community biomass and pielou of biomass, had a positive
relationship with that of species richness (Fig. \@ref(fig:bmr), Table
\@ref(tab:aov), $r_\delta = `r tps_r_sest["p_bm"]`$) while we did not find
significant relationships between the temporal trends of average trophic level,
connectance and the one of species richness (Fig.
\@ref(fig:bmr), Table \@ref(tab:aov)).

@@

We found that species richness, connectance, average trophic level were
spatially positively related to community biomass (resp. 
$r_\delta = `r sp_b_sest["rich"]`$, $`r sp_b_sest["ct"]`$, 
$`r sp_b_sest["tlvl"]`$, Fig.
\@ref(fig:bmr), Table \@ref(tab:aov-sp)) while the pielou of biomass was
negatively related to community biomass ($r_\delta = `r sp_b_sest["p_bm"]`$).

@@

Community biomass, the eveness of biomass were
positively related to species richness (resp. 
$r_\delta = `r sp_r_sest["bm"]`$, $`r sp_r_sest["p_bm"]`$,
$`r sp_r_sest["p_n"]`$) while average
trophic level and connectance were negatively related to
species richness (resp. $r_\delta = `r sp_r_sest["tlvl"]`$, $`r sp_r_sest["ct"]`$,
Fig. \@ref(fig:bmr), Table \@ref(tab:aov-sp)).

@@

We found that the temporal increase of the average trophic level was positively
linked to increase of community biomass in line with a previous model suggesting
that the presence of higher predator can lead to higher community biomass
[@wang_biodiversity_2018]. Increase of average trophic level can be associated
to either gain of biomass in the highest trophic level nor losses of biomass in
the lowest trophic. In our case, such gain or losses of average trophic level is
probably associated to the variations of the presence of large fishes. The
finding that the evenness of biomass, measured by the pielou, decreased through
time as community biomass decreased and reversely, meaning that biomass
distribution across species become less even when community biomass decrease and
reversely. It further indicates that temporal loss of community biomass are
linked with the loss of the higher trophic level.

We also found in space that evenness of biomass was positively related to
community while eveness of abundances had no relation (Fig. SXX). It
constrasts with previous studies which found a negative relationship between
biomass and eveness of abundances in marine and plant communities
[@drobner_relation_1998; @maureaud_biodiversityecosystem_2019] or studies that
found more even plant communities in terms of abundances number display higher
community biomass [@wilsey_biodiversity_2000].

@@

Surprinsingly, we found that the temporal trends of connectance was not related to
the one of species richness as found spatially in most empirical food-web
[@dunne_network_2006] and also in the present study. This result suggests that
the structure of food-webs might be not very sensitive to variation of species
richness and that species loss and gain concerns species of low trophic level,
which in turn does affect connectance very little.

@@

This result has profound implications for conservation. As more and more
studies report community biomass and species richness trends [@hallmann_more_2017;
@ceballos_biological_2017], we documented that they might not be
equivalent in terms of consequences on the structure of the community. In
particular, the loss of the highest trophic level associated to the loss of
community biomass could of have dramatic consequences on ecosystem functioning
as highest predators are prone to stabilise communities
[@shanafelt_stability_2018; danet_species_nodate]. In contrasts, the loss of
species richness was poorly associated to modification of food-web structure,
suggesting that the species loss and gain.

@@

## Conclusion and perspectives

Our study shows that richness and biomass temporal trends in natural fish
communities relate on different ecological processes, biomass temporal dynamics
being related to variation of highest trophic levels while it is not the case for
species richness. Our work also suggests that space for time substitution may be
relevant to predict food-web community structure in case of biomass variation
but not for species richness variation. We look forward to next investigations
to see the generality of assembly rules in food-webs. There are crucial to
better understand the temporal dynamics of food-webs structure and then its
impact on ecosystem functioning.

@@

```{r}
fig_mat_cap <- paste0(
  "(A) Distribution of the fishing sites. (B-F) Temporal trends of biomass,
  species richness, connectance, average trophic level and eveness of biomass
  for two stations."
)
```
```{r, fig.cap=fig_mat_cap, out.width="100%"}
knitr::include_graphics(get_mypath("figures", "matmet.png"))
```

@@

fig1_2 <- get_plot_rich_bm(
  predict_plot = pred_bm_rich_mono_stable_trends,
  get_list = TRUE,
  rm_legend = FALSE,
  bm_x = "log_bm_std",
  bm_y = c("ct_ff", "w_trph_lvl_avg", "piel_bm"), #, "piel_nind"
  rich_y = c("ct_ff", "w_trph_lvl_avg", "log_bm_std", "piel_bm"), #, "piel_nind"
  rich_x = "log_rich_std"
  )
p_bm <- fig1_2$bm
names(p_bm) <- c("ct_ff", "w_trph_lvl_avg", "piel_bm")
#plot_grid(plotlist = p_bm)

@@

```{r}
fig_bm_rich_cap <- paste0(
  "Relationship among the temporal trends of community structure
  and the one of species richness and community biomass. The lines
  surrounding each point represent the standard error associated to the temporal
  trends."
)
```
```{r bmr, out.width = "100%", fig.cap=fig_bm_rich_cap}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~ .x +
    theme(axis.text = element_text(size = 6),
    axis.title = element_text(size = 8)))
fig2$bm <- map(fig2$bm,
  ~ .x + theme(
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 8)
    )
)

plot_grid(
  plot_grid(plotlist = fig2$bm, nrow = 1, labels = c("A", "B", "C")),
  plot_grid(plotlist = fig2$rich, nrow = 1, labels = c("D", "E", "F")),
  nrow = 2
)
```

@@

```{r}
loadd(pred_bm_rich_mono_stable_trends,
  model_bm_rich_int_mono_stable_trends,
  model_bm_rich_int_mono_stable_trends,
  anova_table_bm_rich_mono_stable_trends)

anova_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
  ) %>%
  kable(.,
    booktabs = TRUE,
    row.names = FALSE,
    caption =
      paste0(
      "ANOVA results for the links between the temporal trends of
      community biomass and the temporal trends of the network structure."
    ),
  label = "aov"
  ) %>%
  kable_styling(latex_options = c("scale_down", "stripped")) %>%
  collapse_rows(columns = 1, valign = "top")
```
